LDATS changepoint

### 2019-01-14 paper-comparison/compare-christensen.Rmd
Begin LDA comparison and fix model stash path
start to compare ldas with adjusted/nonadjusted data
store model objects in dropbox

For now I will use the *adjusted rodent table* because this will avoid artificially low abundances for incompletely trapped periods.

```{r switch to adjusted rodent data}

ldats_dat <- as.matrix(ch_dat)
ldats_dat <- apply(ldats_dat, c(1,2), FUN = as.integer)
rodents$document_term_table <- ldats_dat
rm(ldats_dat)
rm(ldats_dates)
rm(ch_dat)
```

@@

```{r save LDAs, eval = F, include = F}
save(ldamodel, file = 'model-stash/paper_lda.Rds')
save(ldats_lda_selected, file = 'model-stash/ldats_lda.Rds')
```

@@

load('model-stash/ldats_lda.Rds')
load('model-stash/paper_lda.Rds')

>>>>

```{r adjusted rodent data}

ldats_dat_adj <- as.matrix(ch_dat)
ldats_dat_adj <- apply(ldats_dat, c(1,2), FUN = as.integer)
rodents_adj <- rodents
rodents_adj$document_term_table <- ldats_dat_adj
rm(ldats_dat_adj)
rm(ch_dat)
```

In principle, whether we use the adjusted or the non-adjusted rodent table should not affect the LDA results. This is because the LDA operates on proportional abundance, not total abundance; a community of abundances 1, 1, 2 would have the same proportions (25/25/50) as a community of abundances 2, 2, 4. However, it's possible that rounding errors could introduce discrepancies between LDAs run on adjusted vs. non-adjusted data.

Exploring/illustrating this:

```{r adj vs non adj LDAS, eval = F}

# paper LDA w adjusted data
source('christensen-ecology-files/AIC_model_selection.R')
source('christensen-ecology-files/LDA-distance.R')

# repeat LDA model fit and AIC calculation with a bunch of different seeds to test robustness of the analysis
best_ntopic = repeat_VEM(rodents_adj[[1]],
                         seeds,
                         topic_min=2,
                         topic_max=6)
hist(best_ntopic$k,breaks=c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5),xlab='best # of topics', main='')

# ==================================================================
# 3. run LDA model
# ==================================================================
ntopics = 4
SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
ldamodel_adj = LDA(rodents[[1]],ntopics, control = list(seed = SEED),method='VEM')

plot(ldamodel_adj)

# paper LDA w nonadjusted data

# ==================================================================
# 3. run LDA model
# ==================================================================
ntopics = 4
SEED = unlist(best_seed)[1]  # For the paper, I use seed 206
ldamodel_nonadj = LDA(rodents[[1]],ntopics, control = list(seed = SEED),method='VEM')

plot(ldamodel_nonadj)

# ldats LDA w adjusted data

ldats_adj_ldas <- LDATS::LDA_set(document_term_table = rodents_adj$document_term_table, topics = c(2:6), nseeds = 100)
ldats_adj_lda_selected <- LDATS::select_LDA(LDA_models = ldats_adj_ldas)

plot(ldats_adj_lda_selected)

# ldats LDA w nonadjusted data

ldats_nonadj_ldas <- LDATS::LDA_set(document_term_table = rodents$document_term_table, topics = c(2:6), nseeds = 100)
ldats_nonadj_lda_selected <- LDATS::select_LDA(LDA_models = ldats_ldas)

plot(ldats_nonadj_lda_selected)

@@

```{r save LDAs, eval = F, include = F}
save(ldamodel, file = '~/Dropbox/ldats-models/paper_lda.Rds')
save(ldats_lda_selected, file = '~/Dropbox/ldats-models/ldats_lda.Rds')
```

@@

load('~/Dropbox/ldats-models/ldats_lda.Rds')
load('~/Dropbox/ldats-models/paper_lda.Rds')

### 2019-01-14 paper-comparison/ldats_changepoint.Rmd
table summarizing cpt outcomes
Begin LDA comparison and fix model stash path
start to compare ldas with adjusted/nonadjusted data
store model objects in dropbox

#### Load LDA ####
load('/Users/renatadiaz/Documents/model-stash/ldats_lda.Rds')
year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

@@

save(ldats_ldats_cpt, file = '/Users/renatadiaz/Documents/model-stash/ldats_lda_ldats_cpt.Rds')
rm(ldats_ldats_cpt)
rm(ldats_lda_selected)
```

```{r load ldats lda + ldats cpt results, echo = F}
load('/Users/renatadiaz/Documents/model-stash/ldats_lda_ldats_cpt.Rds')
```

@@

@@

### Load LDA
```{r load paper LDA}
#### Load paper LDA ####
load('/Users/renatadiaz/Documents/model-stash/paper_lda.Rds')
```

@@

```{r run paper LDA + ldats cpt, eval = F}
paper_ldats_cpt <- TS_on_LDA(LDA_models = ldamodel, document_covariate_table = x, formulas = ~ sin_year + cos_year,nchangepoints = 1:6, weights = NULL)
save(paper_ldats_cpt, file = '/Users/renatadiaz/Documents/model-stash/paper_lda_ldats_cpt.Rds')
```

```{r load paper LDA + ldats cpt, echo = F}
load('/Users/renatadiaz/Documents/model-stash/paper_lda_ldats_cpt.Rds')
```

@@

cpt_dates
```

>>>>

#### Load LDA ####
load('~/Dropbox/ldats-models/ldats_lda.Rds')
year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

@@

save(ldats_ldats_cpt, file = '~/Dropbox/ldats-models/ldats_lda_ldats_cpt.Rds')
rm(ldats_ldats_cpt)
rm(ldats_lda_selected)
```

```{r load ldats lda + ldats cpt results, echo = F}
load('~/Dropbox/ldats-models/ldats_lda_ldats_cpt.Rds')
```

@@

write.csv(cpt_dates, 'ldats_ldats_dates.csv')

@@

### Load LDA
```{r load paper LDA}
#### Load paper LDA ####
load('~/Dropbox/ldats-models/paper_lda.Rds')
```

@@

```{r run paper LDA + ldats cpt, eval = F}
paper_ldats_cpt <- TS_on_LDA(LDA_models = ldamodel, document_covariate_table = x, formulas = ~ sin_year + cos_year,nchangepoints = 1:6, weights = NULL)
save(paper_ldats_cpt, file = '~/Dropbox/ldats-models/paper_lda_ldats_cpt.Rds')
```

```{r load paper LDA + ldats cpt, echo = F}
load('~/Dropbox/ldats-models/paper_lda_ldats_cpt.Rds')
```

@@

write.csv(cpt_dates, 'paper_ldats_dates.csv')
